---
title: "Virtual dissection of MHB region."
output:
  BiocStyle::html_document:
    code_folding: hide
    number_sections: yes
    toc: yes  
  html_document:
    code_folding: hide
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: true
    theme: united
---


# Load dependencies and source data

```{r load_data, message=FALSE}

library(Matrix)
library(scran)
library(BiocParallel)
library(BiocNeighbors)
library(igraph)
library(reshape2)
library(knitr)
library(scater)
library(scales)
library(biomaRt)
library(batchelor)
library(irlba)
library(batchelor)
library(tibble)
library(dplyr)
library(viridis)
library(ggridges)
library(ggpubr)
library(ggforce)
library(binr)
library(tripack)
library(sp)
library(rgdal)
library(maps)
library(ggrepel)
library(wesanderson)
library(plyr)
library(ggsn)

root.dir = "/Users/alsu/Develop/spatial/mouse_embryo/"
figures.dir = paste0(root.dir, "figures/imputation/imputation/brain/")
source(paste0(root.dir, "amScripts_mouseEmbryoFISH/core_functions.R"))
source(paste0(root.dir, "amScripts_mouseEmbryoFISH/visualization_functions.R"))

# load seq
load_embryo_8.5(dir = "local")
sce.seq = sce[!rownames(sce) == "Xist", ]
meta = data.frame(colData(sce.seq))
# add CT
meta.CT = readRDS(paste0(root.dir, "data/8_5/source/celltype_annotation_refined.Rds"))
meta.CT = rownames_to_column(meta.CT, var = "uniqueID")
meta = merge(meta, meta.CT, all.x = T, all.y = F)
CTs.2discard = c("Low quality")
meta = meta[!meta$celltype_mapped_refined %in% CTs.2discard, ]
meta$embryo_z = paste0(meta$embryo, "_", meta$z)

# for brain panel we work with embryo2-2
meta = meta[meta$embryo_z == "embryo2_2",]
sce.seq = sce.seq[, colnames(sce.seq) %in% meta$uniqueID]
counts.seq = assay(sce.seq, "cpm_wo_xist")

# load imputation for the relevant embryo: embryo2, z=2
imputation.embryo2_2 = readRDS(paste0(root.dir, "data/8_5/stitching/imputation/imputation_seq2atlas_logcounts_embryo2z2.rds"))
imputation.embryo2_2 = imputation.embryo2_2[, colnames(imputation.embryo2_2) %in% meta$uniqueID]

# boundary polygons
boundary_polygons = getSegmentationVerticesDF(colData(sce.seq),
                                              xname = "segmentation_vertices_x_global_affine",
                                              yname = "segmentation_vertices_y_global_affine",
                                              othercols = c("uniqueID", "embryo", "z"))
boundary_polygons = merge(boundary_polygons, meta[, c("uniqueID", "celltype_mapped_refined")], all.x = F, all.y = T)



```


# Otx2 & Gbx2 expression - all brain

Region for virtual dissection is selected manually.

Supplementary figures 19 D,E.

```{r Otx2-Gbx2, warning=FALSE}

counts.gbx2 = counts.seq[rownames(counts.seq) == "Gbx2", ]
counts.gbx2 = data.frame(uniqueID = names(counts.gbx2), counts.gbx2 = counts.gbx2)
counts.otx2 = counts.seq[rownames(counts.seq) == "Otx2", ]
counts.otx2 = data.frame(uniqueID = names(counts.otx2), counts.otx2 = counts.otx2)
current.counts = merge(counts.gbx2, counts.otx2)
current.counts = join(current.counts, boundary_polygons, type = "right")
current.counts = current.counts[current.counts$celltype_mapped_refined == "Forebrain/Midbrain/Hindbrain" , ]

zoomedRegion.coords = data.frame(x1 = -0.5 , x2 = 1.15 , y1 = 2 , y2 = 3.15)

# Gbx2
current.counts = current.counts[order(current.counts$counts.gbx2),]
p1 <- ggplot() + 
  geom_polygon(data = current.counts, 
               aes(x = segmentation_vertices_x_global_affine, y = -segmentation_vertices_y_global_affine, group = uniqueID, fill = counts.gbx2)) + 
  scale_fill_gradient2(low = "gray75", high = "black", mid = "cornflowerblue", midpoint = (max(current.counts$counts.gbx2) + min(current.counts$counts.gbx2))/2, name = "") + guides(fill = F) +
  ylim(c(0,3.4)) +
  ggtitle("Gbx2") +
  geom_rect(data = zoomedRegion.coords, mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2), color="black", alpha=0) + 
  labs(x="",y="") +
  coord_fixed() + 
  theme_classic() + 
  add_scalebar(x = 2.75 , y = 0) + 
  theme(text = element_text(size=23)) +  
  theme(line = element_blank(), axis.text.x = element_blank(), axis.text.y = element_blank())
p1


# Otx2
current.counts = current.counts[order(current.counts$counts.otx2),]
p2 <- ggplot() + 
  geom_polygon(data = current.counts, 
               aes(x = segmentation_vertices_x_global_affine, y = -segmentation_vertices_y_global_affine, group = uniqueID, fill = counts.otx2)) + 
  scale_fill_gradient2(low = "gray75", high = "black", mid = "cornflowerblue", midpoint = (max(current.counts$counts.otx2) + min(current.counts$counts.otx2))/2, name = "") + guides(fill = F) +
  ylim(c(0,3.4)) +
  ggtitle("Otx2") +
  geom_rect(data = zoomedRegion.coords, mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2), color="black", alpha=0) + 
  labs(x="",y="") +
  coord_fixed() + 
  theme_classic() + 
  add_scalebar(x = 2.75 , y = 0) + 
  theme(text = element_text(size=23)) +  
  theme(line = element_blank(), axis.text.x = element_blank(), axis.text.y = element_blank())

  
p <- ggarrange(p1,p2,ncol=2)
p
ggsave(filename = paste0(figures.dir, "spatial_otx2_gbx2.png"), plot = p, width = 8, height = 4)




```

## Defining MHB 

Midbrain-hindbrain border is selected manually based on expression of Otx2 and Gbx2.

Supplementary figure 19 H.

```{r get-mhb, warning=FALSE}


meta.section = meta[meta$x_global_affine > -0.5 & meta$x_global_affine < 1.15 &
                                          -1*meta$y_global_affine > 2 & -1*meta$y_global_affine < 3.15 &
                                          meta$celltype_mapped_refined == "Forebrain/Midbrain/Hindbrain", ]
meta.section$y_global_affine_rev = -1*meta.section$y_global_affine


# dissection region (manually selected)
anchor.points.x = c(0.12, 0.35, 0.42, 0.45, 0.45, -0.51, -1, -1, -0.51, 0)
anchor.points.y = c(3.05, 2.7, 2.5, 2.25, 1.99, 1.99, 2, 3.26, 3.26, 3.26)
anchor.points = data.frame(x_global_affine = anchor.points.x, y_global_affine_rev = anchor.points.y)


group_colours = c("Brain" = "firebrick1", "0" = "gray80", "anchor" = "black")
p <- ggplot() +
  geom_point(data = meta.section, aes(x = x_global_affine, y = -y_global_affine) , size = 2.5, alpha = .75) +
  geom_polygon(data = anchor.points, aes(x = x_global_affine, y = y_global_affine_rev), col = "firebrick2", fill = NA, size = 2) +
  theme(text = element_text(size=10)) + 
  scale_x_continuous(breaks = seq(-2.5, 1, by = 0.5)) + 
  scale_y_continuous(breaks = seq(-.5, 3, by = 0.5)) 
p


# split in M and H
subgroup = sapply(1:nrow(meta.section), function(i){
  x = meta.section$x_global_affine[i]
  y = meta.section$y_global_affine_rev[i]
  if (point.in.polygon(x, y, anchor.points$x_global_affine, anchor.points$y_global_affine_rev) == 1){
    return("Hindbrain")
  } else if ( x > 0.9 & y < 2.13 ){
    return("Midbrain")
  } else {
    return("Midbrain")
  }
})
meta.section$subgroup = subgroup


meta.section.boundaries = join(boundary_polygons, meta.section[,c("uniqueID", "subgroup")], type = "right")
subgroup_colours = c("Midbrain" = "coral", "Hindbrain" = "darkorchid4")
p <- ggplot() + 
  geom_polygon(data = meta.section.boundaries,
               aes(x = segmentation_vertices_x_global_affine, y = -segmentation_vertices_y_global_affine, group = uniqueID, fill = subgroup)) + labs(x = "", y = "") + 
  scale_fill_manual(values = subgroup_colours, name = "Region") + 
  coord_fixed() +
  theme_classic() +
  add_scalebar(x = 1.75 , y = 1.75) +
  theme(text = element_text(size=23)) + 
  theme(line = element_blank(), axis.text.x = element_blank(), axis.text.y = element_blank())
p
ggsave(filename = paste0(figures.dir, "virtual_dissection__MHB.png"), plot = p, width = 7, height = 4)



```

# Otx2 - Gbx2 -- zoomed

Supplementary figure 19 F,G; I.

```{r Otx2-Gbx2-zoomed-in, warning=FALSE}


counts.gbx2 = counts.seq[rownames(counts.seq) == "Gbx2", ]
counts.gbx2 = data.frame(uniqueID = names(counts.gbx2), counts.gbx2 = counts.gbx2)
counts.otx2 = counts.seq[rownames(counts.seq) == "Otx2", ]
counts.otx2 = data.frame(uniqueID = names(counts.otx2), counts.otx2 = counts.otx2)
current.counts = merge(counts.gbx2, counts.otx2)
current.counts = join(current.counts, boundary_polygons, type = "right")
current.counts = join(current.counts , meta.section[, c("uniqueID", "x_global_affine", "y_global_affine")] , type = "right")


anchor.points.x = c(0.12, 0.35, 0.42, 0.45, 0.45, 0)
anchor.points.y = c(3.05, 2.7, 2.5, 2.25, 1.99, 3.26)
anchor.points = data.frame(x_global_affine = anchor.points.x, y_global_affine_rev = anchor.points.y)


# spatial
current.counts = current.counts[order(current.counts$counts.gbx2),]
p1 <- ggplot() + 
  geom_polygon(data = current.counts,
           aes(x = segmentation_vertices_x_global_affine, y = -segmentation_vertices_y_global_affine, group = uniqueID, fill = counts.gbx2)) + 
  scale_fill_gradient2(low = "gray75", high = "black", mid = "cornflowerblue", midpoint = (max(current.counts$counts.gbx2) + min(current.counts$counts.gbx2))/2, name = "") + guides(fill = F) +
  labs(x = "x", y = "y") +
  ggtitle("Gbx2") +
  labs(x="",y="") +
  coord_fixed() +
  geom_line(data = anchor.points, aes(x = x_global_affine, y = y_global_affine_rev), color = "firebrick4", size = 1.5) +
  theme_classic() + add_scalebar(x = 1.75 , y = 1.75) + 
  theme(text = element_text(size=23)) +  
  theme(line = element_blank(), axis.text.x = element_blank(), axis.text.y = element_blank()) 
  
  
current.counts = current.counts[order(current.counts$counts.otx2),]
p2 <-  ggplot() + 
  geom_polygon(data = current.counts,
           aes(x = segmentation_vertices_x_global_affine, y = -segmentation_vertices_y_global_affine, group = uniqueID, fill = counts.otx2)) + 
  scale_fill_gradient2(low = "gray75", high = "black", mid = "cornflowerblue", midpoint = (max(current.counts$counts.otx2) + min(current.counts$counts.otx2))/2, name = "") + guides(fill = F) +
  labs(x = "x", y = "y") +
  ggtitle("Otx2") +
  labs(x="",y="") +
  coord_fixed() +
  geom_line(data = anchor.points, aes(x = x_global_affine, y = y_global_affine_rev), color = "firebrick4", size = 1.5) +
  theme_classic() + add_scalebar(x = 1.75 , y = 1.75) +
  theme(text = element_text(size=23)) +  
  theme(line = element_blank(), axis.text.x = element_blank(), axis.text.y = element_blank())
p <- ggarrange(p1,p2, ncol=2)  
p
ggsave(filename = paste0(figures.dir, "spatial_otx2_gbx2__zoomed.png"), plot = p, width = 8, height = 4)



# violins 
current.counts.grouped <- as.data.frame( current.counts %>% group_by(uniqueID) %>% 
                                  dplyr::summarise(counts.gbx2 = mean(counts.gbx2) , 
                                                   counts.otx2 = mean(counts.otx2) ))
current.counts.grouped = merge(current.counts.grouped, meta.section[, c("uniqueID", "subgroup")])
subgroup_colours = c("Midbrain" = "coral", "Hindbrain" = "darkorchid4")

p1 <- ggplot(current.counts.grouped, aes(x = subgroup , y = counts.gbx2, fill = subgroup)) + 
  geom_violin(aes(fill = subgroup), alpha = .5, color = "transparent") + 
  scale_fill_manual(values = subgroup_colours,  name = "") +
  geom_boxplot(outlier.alpha = 0, coef = 0, color = "gray40", width = 0.1) +
  labs(x = "", y = "Gbx2 counts") + 
  theme_classic() + theme(text = element_text(size=16)) + theme(axis.text.x = element_blank()) 
p2 <- ggplot(current.counts.grouped, aes(x = subgroup , y = counts.otx2, fill = subgroup)) + 
  geom_violin(aes(fill = subgroup), alpha = .5, color = "transparent") + 
  scale_fill_manual(values = subgroup_colours, name = "") +
  geom_boxplot(outlier.alpha = 0, coef = 0, color = "gray40", width = 0.1) +
  labs(x = "", y = "Otx2 counts") + 
  theme_classic() + theme(text = element_text(size=16)) + theme(axis.text.x = element_blank()) 
  
  
p <- ggarrange(p1,p2, common.legend = T, ncol=2)
p
ggsave(filename = paste0(figures.dir, "violin__Otx2_Gbx2.png"), plot = p, width = 8, height = 4)



```

## DE - between M and H

Figure 3E.

```{r DE, warning=FALSE}


library(wesanderson)
pal = wes_palette("Darjeeling1" , 2)
FDR.thresh = 0.05
lfc.thresh = .2
imputation.embryo2_2_section = imputation.embryo2_2[, colnames(imputation.embryo2_2) %in% meta.section$uniqueID]
sce.section = SingleCellExperiment(list(logcounts=imputation.embryo2_2_section))

de.section = findMarkers(sce.section, meta.section$subgroup, lfc = lfc.thresh)
de.section = as.data.frame(de.section[[2]])
de.section = rownames_to_column(de.section, var = "Gene")
de.section = de.section[, c("Gene", "FDR", "logFC.Hindbrain")]
colnames(de.section) = c("Gene", "padj", "log2FoldChange")

# add average expression
counts.section = logcounts(sce.section)
counts.mean.section = apply(counts.section, 1, mean)
counts.mean.section = data.frame(Gene = names(counts.mean.section), baseMean = counts.mean.section)
de.section = merge(de.section, counts.mean.section, by = "Gene")

de.section.sign = de.section[de.section$padj <= FDR.thresh , ]
write.table(de.section.sign, file = paste0(root.dir, "data/sup_tables/deGenes_MHB.tab"), row.names = F, col.names = T, quote = F, sep = "\t")


significance = sapply(1:nrow(de.section), function(i){
  if (de.section$padj[i] < FDR.thresh & de.section$log2FoldChange[i] > 0){
    return("Up")
  } else if(de.section$padj[i] < FDR.thresh & de.section$log2FoldChange[i] < 0 ){
    return("Down")
  } else {
    return("0")
  }
})
de.section$significance = significance              
de.section$significance = factor(de.section$significance)
de.section = de.section[order(de.section$significance) , ]
pal = wes_palette("Darjeeling1" , 2)      
subgroup_colours = c("Midbrain" = "coral", "Hindbrain" = "darkorchid4", "Forebrain" = "darkslategray3")
significance_colors = c("Up" = "coral", "Down" = "darkorchid4", "0" = "gray75")                     
                    

lfc.thresh_strong = 0.5 
p <- ggplot(de.section, aes(x = baseMean, y = log2FoldChange, color = significance)) + 
  geom_point() + 
  theme_classic() +
  scale_colour_manual(values = significance_colors) +
  theme(legend.position = "none")+
  geom_hline(yintercept = 0, col = "black", linetype = "dotted", size = 1) +
  geom_text_repel(data=subset(de.section,Gene %in% de.section$Gene[de.section$significance == "Up" & de.section$log2FoldChange > lfc.thresh_strong]),
                     aes(baseMean, log2FoldChange, label = Gene),size = 4, color="black") + 
  geom_text_repel(data=subset(de.section,Gene %in% c("Gbx2", de.section$Gene[de.section$significance == "Down" & de.section$log2FoldChange < -1*lfc.thresh_strong])),
                     aes(baseMean, log2FoldChange, label = Gene),size = 4, color="black") +
  labs(x = "average logcounts", y = "log2 Fold Change") + 
  theme(text = element_text(size=16)) + 
  ggtitle(paste0("Up: ", sum(de.section$significance == "Up"), ", Down: ", sum(de.section$significance == "Down")))
p
ggsave(filename = paste0(figures.dir, "de_MA_V3.png"), plot = p, width = 7, height = 4)


```


# BULK

## Spatial -- DE genes

```{r spatial-DE, warning=FALSE}


# Section
getSpatialPlot = function(gene, subgroup){
  current.counts.imp = imputation.embryo2_2[rownames(imputation.embryo2_2) == gene, ]
  current.counts.imp = data.frame(uniqueID = names(current.counts.imp), counts.imp = current.counts.imp)
  current.counts = merge(current.counts.imp, meta.section.boundaries, all.y = T, all.x = F)
  
  # plot
  current.counts = current.counts[order(current.counts$counts.imp),]
  p <- ggplot(current.counts,
           aes(x = segmentation_vertices_x_global_affine, y = -segmentation_vertices_y_global_affine, group = uniqueID, fill = counts.imp)) + 
  geom_polygon() + 
  scale_fill_gradient2(low = "gray75", high = "black", mid = "cornflowerblue", midpoint = ( max(current.counts$counts.imp) + min(current.counts$counts.imp))/2, name = "") +
  guides(fill = F) +
  labs(x = "", y = "") +
  #ggtitle(gene) +
  coord_fixed() + 
  theme_classic() +
  theme(text = element_text(size=23)) + 
  theme(line = element_blank(), axis.text.x = element_blank(), axis.text.y = element_blank())
  ggsave(filename = paste0(figures.dir, "spatial/", subgroup, "/section/", gene, ".png"), plot = p, width = 5, height = 4)
  return(p)
}

genes.hindbrain = de.section$Gene[de.section$logFC.Hindbrain < -0.25]
for (current.gene in genes.hindbrain){
  getSpatialPlot(current.gene, "hindbrain")
}
genes.midbrain = de.section$Gene[de.section$logFC.Hindbrain > 0.25]
for (current.gene in genes.midbrain){
  getSpatialPlot(current.gene, "midbrain")
}


# Whole brain
getSpatialPlot = function(gene, subgroup){
  current.counts.imp = imputation.embryo2_2[rownames(imputation.embryo2_2) == gene, ]
  current.counts.imp = data.frame(uniqueID = names(current.counts.imp), counts.imp = current.counts.imp)
  current.counts = merge(current.counts.imp, boundary_polygons, all.y = T, all.x = F)
  current.counts = current.counts[current.counts$celltype_mapped_refined == "Forebrain/Midbrain/Hindbrain",]
  current.counts$x1 = -0.5
  current.counts$x2 = 1.15
  current.counts$y1 = 2
  current.counts$y2 = 3.15
  # plot
  current.counts = current.counts[order(current.counts$counts.imp),]
  p <- ggplot(current.counts,
           aes(x = segmentation_vertices_x_global_affine, y = -segmentation_vertices_y_global_affine, group = uniqueID, fill = counts.imp)) + 
  geom_polygon() + 
  scale_fill_gradient2(low = "gray75", high = "black", mid = "cornflowerblue", midpoint = ( max(current.counts$counts.imp) + min(current.counts$counts.imp))/2, name = "") +
  guides(fill = F) +
  labs(x = "", y = "") +
  geom_rect(mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2), color="black", alpha=0) + 
  #ggtitle(gene) +
  coord_fixed() + 
  theme_classic() +
  theme(text = element_text(size=23)) + 
  ylim(c(0,3.4)) + 
  theme(line = element_blank(), axis.text.x = element_blank(), axis.text.y = element_blank())
  ggsave(filename = paste0(figures.dir, "spatial/", subgroup, "/wholeBrain/", gene, ".png"), plot = p, width = 5, height = 4)
  return(p)
}

genes.hindbrain = de.section$Gene[de.section$log2FoldChange < -0.25]
for (current.gene in genes.hindbrain){
  getSpatialPlot(current.gene, "hindbrain")
}
genes.midbrain = de.section$Gene[de.section$log2FoldChange > 0.25]
for (current.gene in genes.midbrain){
  getSpatialPlot(current.gene, "midbrain")
}









```

#Session Info

```{r sessinf}

sessionInfo()

```
