---
title: "smFISH genes: assessment of the dependance between intensties and channels; hyb rounds."
output:
  BiocStyle::html_document:
    code_folding: hide
    number_sections: yes
    toc: yes  
  html_document:
    code_folding: hide
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: true
    theme: united
---

# Load dependencies and source data

Look into which parameters affect manual threshold.

**Gene = channelLabel + folderLabel.**
**ID = position + gene.** 


```{r load_data, message=FALSE}

library(Matrix)
library(scran)
library(igraph)
library(reshape2)
library(knitr)
library(scater)
library(scales)
library(biomaRt)
library(VennDiagram)
library(RColorBrewer)
library(tibble)
library(ggplot2)
library(tidyverse)
library(ggpubr)
library(ggridges)
library(viridis)
library(MASS)
library(mclust, quietly=TRUE)
library(sfsmisc) 
library(irlba)
library(caret)
library(MASS)
library(ggforce)
library(wesanderson)
set.seed(32)


data.dir = "/Users/alsu/Develop/spatial/mouse_embryo/data/8_5/smFISH_thresholds/"
figures.dir = "/Users/alsu/Develop/spatial/mouse_embryo/figures/smFISH__dotsAutoDetect/"

# load data 
manual_thresh.embryo_12 = read.table(paste0( data.dir, "manual_threshold/embryo12.tab"), header = TRUE, sep = "\t", stringsAsFactors = FALSE, comment.char = "$")
manual_thresh.embryo_12$gene = paste0(manual_thresh.embryo_12$channelLabel, "_", manual_thresh.embryo_12$folderLabel)
manual_thresh.embryo_12$id = paste0(manual_thresh.embryo_12$gene, "_", manual_thresh.embryo_12$position)


manual_thresh.embryo_3 = read.table(paste0( data.dir, "manual_threshold/embryo3.tab"), header = TRUE, sep = "\t", stringsAsFactors = FALSE, comment.char = "$")
manual_thresh.embryo_3$gene = paste0(manual_thresh.embryo_3$channelLabel, "_", manual_thresh.embryo_3$folderLabel)
manual_thresh.embryo_3$id = paste0(manual_thresh.embryo_3$gene, "_", manual_thresh.embryo_3$position)


# channels that are affected in the first experiment: thresholds for these FOVs were re-calculated again
manual_thresh.embryo_12_weird = read.table(paste0( data.dir, "manual_threshold/embryo12_channel_1_pos_39_40_43_44.tab"), header = TRUE, sep = "\t", stringsAsFactors = FALSE, comment.char = "$")
manual_thresh.embryo_12_weird$gene = paste0(manual_thresh.embryo_12_weird$channelLabel, "_", manual_thresh.embryo_12_weird$folderLabel)
manual_thresh.embryo_12_weird$id = paste0(manual_thresh.embryo_12_weird$gene, "_", manual_thresh.embryo_12_weird$position)


```


## Position-Channel-Folder VS manually selected thresholds

Let's see if there is any pattern/dependancy between position-channel-folder and manually selected thresholds.

Supplementary figure 24 A, B.

```{r get-stat, fig.cap="Manual thresholding VS position-channel-folder.", fig.height=10, fig.wide=TRUE, warning=FALSE}


get_plot = function(stat){
  p1 <- ggplot(stat, aes(x = factor(channelLabel), y = manual.thresh_upper)) + 
  geom_violin(color = "firebrick4", fill = "gray90", alpha = 0.5) +
  geom_sina(color = "firebrick4", alpha = .75, size=3) +
  labs(x = "Channel", y = "Intensity threshold") +
  theme_classic() +
  theme(legend.position = "none") + theme(text = element_text(size=16))
p2 <- ggplot(stat, aes(x = factor(folderLabel), y = manual.thresh_upper)) + 
  geom_violin(color = "forestgreen", fill = "gray90", alpha = 0.5) +
  geom_sina(color = "forestgreen", alpha = .75, size=3) +
  labs(x = "Hybridization round", y = "Intensity threshold") +
  theme_classic() +
  theme(legend.position = "none") + theme(text = element_text(size=16))

  p <- ggarrange(p1,p2,ncol=2, nrow=1)
  return(p)
}

p = get_plot(manual_thresh.embryo_12)
ggsave(filename = paste0(figures.dir, "stat/embryo12_parameters_VS_manualThreshold", ".png"), plot = p, width = 12, height = 10)
 
p = get_plot(manual_thresh.embryo_3)
ggsave(filename = paste0(figures.dir, "stat/embryo3_parameters_VS_manualThreshold", ".png"), plot = p, width = 12, height = 10)
 

```

**There is a dependancy between channel and threshold. **

## Get intensity thresholds for all genes, FOVs.

```{r get-intensity-thresh, fig.cap="Manual thresholding VS position-channel-folder.", fig.height=10, fig.wide=TRUE, warning=FALSE}

unq.positions = c(0:50)
unq.channels = c(1:3)
unq.folders = c(1:12)

# embryos 1 and 2
stat.batch_1 = lapply(unq.positions, function(position){
  stat.position = lapply(unq.channels, function(channel){
    stat.channel = lapply(unq.folders, function(folder){
      gene = paste0(channel, "_", folder)
      id = paste0(channel, "_", folder, "_", position)
      idx = which(manual_thresh.embryo_12_weird$id == id )
      if (length(idx) == 1){
        out = data.frame(position = position, channelLabel = channel, folderLabel = folder, id = id, gene = gene,
                         thresh = manual_thresh.embryo_12_weird$manual_upper[idx])
      } else {
        idx = which(manual_thresh.embryo_12$channelLabel == channel )
        out = data.frame(position = position, channelLabel = channel, folderLabel = folder, id = id, gene = gene,
                         thresh = round( mean( manual_thresh.embryo_12$manual.thresh_upper[idx]) ) )
      }
      return(out)
    })
    stat.channel = do.call(rbind, stat.channel)
    return(stat.channel)
  })
  stat.position = do.call(rbind, stat.position)
  return(stat.position)
})
stat.batch_1 = do.call(rbind, stat.batch_1)
write.table(stat.batch_1 , file = paste0(data.dir, "mean_prediction_embryo12.tab"), row.names = F, col.names = T, sep = "\t") 


# embryo3
unq.positions = c(0:39)
unq.channels = c(1:3)
unq.folders = c(1:12)

stat.batch_2 = lapply(unq.positions, function(position){
  stat.position = lapply(unq.channels, function(channel){
    stat.channel = lapply(unq.folders, function(folder){
      gene = paste0(channel, "_", folder)
      id = paste0(channel, "_", folder, "_", position)
      idx = which(manual_thresh.embryo_3$channelLabel == channel )
      out = data.frame(position = position, channelLabel = channel, folderLabel = folder, id = id, gene = gene,
                         thresh = round( mean( manual_thresh.embryo_3$manual.thresh_upper[idx]) ) )
      return(out)
    })
    stat.channel = do.call(rbind, stat.channel)
    return(stat.channel)
  })
  stat.position = do.call(rbind, stat.position)
  return(stat.position)
})
stat.batch_2 = do.call(rbind, stat.batch_2)
write.table(stat.batch_2 , file = paste0(data.dir, "mean_prediction_embryo3.tab"), row.names = F, col.names = T, sep = "\t") 




```

# Session Info

```{r sessinf}
sessionInfo()
```
